
### Instructions

- Windows 11
    This code is designed for RStudio on Windows 11. It *may* run on Windows 10.

- Admin Execution
    The code is designed for execution with administrator privileges, i.e.
    RStudio was invoked via the Windows "Run as administrator" option.

- Ghostscript
    It's assumed Ghostscript is available on the system, and an invocation path
    has been set in the RStudio environment variables "GS_CMD" and "R_GSCMD",
    i.e., the path to a relevant executable such as "gswin64c.exe".
    The code was not designed for use with Ghostscript more recent than version
    10.00.0.

- PowerShell
    It's assumed PowerShell is available on the system. It's typically installed
    by default in Windows 10 and 11.

- Stan
     It's assumed R and RStudio have been configured to enable execution of
     RStan code (Stan's R interface). For more information, see:
     https://mc-stan.org/rstan/

- ThrottleStop
    The code is designed for use with ThrottleStop. The code will however run
    without it.

- Task Scheduler
    The code toggles activation of the Windows Task Scheduler via dynamic
    registry manipulation. Full implementation (either full shutdown or full
    startup) requires also restarting Windows.

- Run Sequence
    The intended sequence of execution is as follows:
     1. Run chunk "setup".
     2. Run chunk "mcmc_sampling".
     3. Run chunk "test_convergence". It may be necessary to modify the number
          of warmup and non-warmup iterations of the Stan sampler, to achieve
          convergence on your system.
     4. Run chunk "dampen_os_noise".
     5. Save the RStudio workspace; from the RStudio main menu, select:
          "Session" --> "Save Workspace As ...".
     6. Restart Windows.
     7. Restart RStudio and load the session saved at step 5, i.e., from the
          RStudio main menu, select:
            "Session" --> "Load Workspace".
     8. Start ThrottleStop, and check its "Disable Turbo" checkbox.
     9. Run the chunk "test_mcmc_throttled".
    10. Close ThrottleStop.
    11. Run the chunk "test_mcmc_unthrottled".
    12. Save final results via "Session" --> "Save Workspace As ..." from the
          RStudio main menu.
    13. Restart Windows.


```{r setup, include=FALSE}

library(extrafont)      # To change font styles in ggplot2.
library(grDevices)      # To produce advanced plots.
library(grid)           # To produce advanced plots.
library(knitr)          # To set chunk options.
library(latex2exp)      # To add LaTeX to plots.
library(parallel)       # To detect available processing cores.
library(posterior)      # For advanced MCMC diagnostics.

library(ggplot2)        # An egg dependency.
library(gridExtra)      # An egg dependency.
library(egg)            # To produce advanced plots.

library(StanHeaders)    # An rstan dependency.
library(rstan)

# Global knitr options.
opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)

# Avoid recompilation of unchanged Stan programs.
rstan_options(auto_write = TRUE)

# Notionally global "constant".
shape_round <- 16

# Configure number of cores and chains to use.
# N.B. Using 'logical = FALSE' detects physical cores.
#      Using 'logical = TRUE' (the default) detects logical cores.
hyperthreading_status <- TRUE
active_cores <- detectCores(logical = hyperthreading_status)
options(mc.cores = active_cores)


# Utility functions.

background_services_disable <- function() {

  # Dynamically disable superfluous system services and processes.
  #
  # Returns,
  #   a list of lists: list(services = list_service, processes = list_process),
  #
  #   where,
  #   list_service    a list describing services that were shut down.
  #   list_process    a list describing processes that were shut down.
  #
  #   It's intended this list of lists be used to restart the services and
  #   processes shut down here.
  #
  #   N.B. If the function doesn't complete, NULL is returned.


  # Some services and processes take a few minute to complete startup.
  # Ensure the OS session has been running for at least five minutes.
  last_logon <- sys::as_text(
    sys::exec_internal(
      "powershell",
      args = paste0("net user ", Sys.info()[6],
                    " | findstr /B /C:'Last logon'")
    )$stdout
  )
  last_logon <- strptime(
    stringr::str_trim(stringr::str_remove(last_logon, "Last logon")),
    "%d/%m/%Y %I:%M:%S %p"
  )
  time_diff <- as.integer(difftime(Sys.time(), last_logon, units = "mins"))
  if (time_diff < 5) {
    cat("Function cancelled. Cannot begin < five minutes after system logon.")
    mins_left <- 5 - floor(time_diff)
    if (mins_left == 1) {
      str_suffix <- " minute."
    } else {
      str_suffix <- " minutes."
    }
    cat("\nTry again in ", as.character(english::as.english(mins_left)),
        str_suffix, sep = "")
    return(NULL)
  }
  rm(last_logon)
  rm(time_diff)


  # Ensure the session is being run with administrator privileges.
  # N.B. Test for text "True" / "False" here, not Boolean TRUE / FALSE.
  if (
    sys::as_text(
      sys::exec_internal(
        "powershell",
        args = "[Security.Principal.WindowsIdentity]::GetCurrent().Groups -contains 'S-1-5-32-544'"
      )$stdout
    ) != "True"
  ) {
    cat("Function cancelled. Should not be executed without elevated rights.")
    return(NULL)
  }


  # Shut down the Task Scheduler.
  # Modify the registry key:
  #   HKEY_LOCAL_MACHINE\SYSTEM\CurrentControlSet\Services\Schedule.
  #
  # N.B. Change the key's Start value. Assume its original value is 2.
  #        (Here 2 is an alias for "Automatic (Delayed Start)")
  #        (Here 4 is an alias for "Disabled)
  #
  # Once complete, the Schedule service should appear as Stopped.
  #
  print("Disabling Task Scheduler...")

  # Does the registry key need updating?
  key_value <- sys::as_text(
    sys::exec_internal(
      "powershell",
      args = "$key_val = Get-ItemProperty -Path 'HKLM:\\SYSTEM\\CurrentControlSet\\Services\\Schedule'; $key_val.Start"
    )$stdout
  )
  if (as.integer(key_value) == 2) {
    sys::exec_wait(
      "powershell",
      args = "Set-ItemProperty -Path 'HKLM:\\SYSTEM\\CurrentControlSet\\Services\\Schedule' -Name 'Start' -Value 4"
    )
    cat("Task Scheduler registry details updated. Please restart system.")
    return(NULL)
  }

  # Is the service still running?
  running_test <- tryCatch(
    expr = {
      sys::as_text(
        sys::exec_internal(
          "powershell",
          args = c(
            "Get-Service -Name 'Schedule' | Where-Object {$_.Status -eq 'Running'}"
          )
        )$stdout
      )
    },
    error = function(e) {
      NULL
    }
  )
  if (!is.null(running_test)) {
    if (length(running_test) > 0) {
      cat("Task Scheduler set to shut down. Please restart system.")
      return(NULL)
    }
  }


  # Processes section ##########################################################

  print("Disabling processes...")

  # Set the processes to be stopped.
  # Successfully disabling Adobe processes is, prima facie, sensitive to shut
  # down order.

  consolidated_process_names <- NULL
  target_processes <- c(

    # Creative Cloud.
    "Adobe` Desktop` Service",

    # Provides dynamic content for Adobe Creative Cloud.
    "CCXProcess",

    # Syncs data between a computer and Creative Cloud account.
    "CoreSync",

    # Facilitates sign-in to Adobe Creative Cloud apps.
    "Creative` Cloud` Helper",

    # Node.js JavaScript Runtime (used by Creative Cloud).
    "node",

    # Adobe Collaboration Synchroniser.
    "AdobeCollabSync",

    # Notification manager for Adobe Creative Cloud.
    "AdobeNotificationClient",

    # Adobe Acrobat Update Service.
    "armsvc",

    # Adobe AcroTray.
    "acrotray",

    # Notification manager for Adobe Acrobat.
    "AcrobatNotificationClient",

    # Facilitates communications between Adobe apps.
    "AdobeIPCBroker",

    # Changes audio settings and converts audio file formats.
    "HPAudioSwitch"
  )

  list_process <- list(
    names = target_processes,
    exists = rep(FALSE, times = length(target_processes)),
    paths = rep(NULL, times = length(target_processes)),
    was_suspended = rep(FALSE, times = length(target_processes))
  )

  for (i in seq_along(list_process$names)) {

    # Test if processes exist via PowerShell.
    exists_test <- tryCatch(
      expr = {
        sys::exec_internal(
          "powershell",
          args = paste0(
            "Get-Process '", list_process$names[i],
            "' -ErrorAction SilentlyContinue"
          )
        )
      },
      error = function(e) {
        NULL
      }
    )

    if (!is.null(exists_test)) {
      list_process$exists[i] <- TRUE

      # Test if processes are suspended.
      suspended_test <- tryCatch(
        expr = {
          sys::exec_internal(
            "powershell",
            args = paste0(
              "$process = Get-Process '", list_process$names[i], "'
              foreach ($thread in $process.Threads) {
                if ($thread.ThreadState -eq 'Wait') {
                  if ($thread.WaitReason.ToString() -eq 'Suspended') {
                    # Deliberately throw an error to signify a process with
                    # suspended threads.
                    throw
                  }
                }
              }"
            )
          )
        },
        error = function(e) {
          NULL
        }
      )
      if (is.null(suspended_test)) {
        list_process$was_suspended[i] <- TRUE
      }

      # Build process names string.
      if (is.null(consolidated_process_names)) {
        consolidated_process_names <- paste0(list_process$names[i])
      } else {
        consolidated_process_names <- paste0(
          consolidated_process_names, ", ", list_process$names[i]
        )
      }
      #
    }
  }

  if (!is.null(consolidated_process_names)) {

    # Split process names into individual strings.
    tmp_pcs <- paste0(
      "'",
      stringr::str_replace_all(
        stringr::str_remove_all(consolidated_process_names, "`"),
        ", ",
        "', '"
      ),
      "'"
    )
    tmp_paths <- paste(
      rep(
        "''", length(strsplit(consolidated_process_names, split = ", ")[[1]])
      ),
      collapse = ", "
    )

    # Get process paths.
    list_process$paths <- sys::as_text(
      sys::exec_internal(
        "powershell",
        args = paste0(
          "$index = 0\n$paths = @(",
          tmp_paths,
          ")\n$processes = @(",
          tmp_pcs,
          ")\nforeach ($process in $processes) {$tmpvar = Get-Process $process -ErrorAction SilentlyContinue\nif ( $tmpvar -is [array] ) {$tmppath = $tmpvar.Path[1]} else {$tmppath = $tmpvar.Path}\n$paths[$index] = $tmppath\n$index = $index + 1}\n$paths"
        )
      )$stdout
    )

    # Close relevant processes.
    sys::exec_wait(
      "powershell",
      args = paste0(
        "Stop-Process -Name ", consolidated_process_names, " -Force"
      )
    )
    # Alternative for sessions not being run with elevated rights.
    # Currently disabled.
    # sys::exec_wait(
    #   "powershell",
    #   args = paste0(
    #     "Start-Process PowerShell -ArgumentList '-Command & {Stop-Process -Name ",
    #     consolidated_process_names, " -Force}' -Verb RunAs -WindowStyle hidden"
    #   )
    # )
  }


  # Services section ###########################################################

  print("Disabling services....")

  # Set the services to be stopped.
  consolidated_service_names <- NULL
  target_services <- c(

    "AdobeARMservice",

    "AdobeUpdateService",

    # Adobe Genuine Software Service.
    "AGMService",

    # Adobe Genuine Software Integrity Service.
    "AGSService",

    # Bluetooth audio device / wireless headphones service.
    "BthAvctpSvc",

    # Automated checking for Office updates.
    "ClickToRunSvc",

    # Intel Content Protection HECI Service. Plays certain premium video types.
    "cphs",

    # Prevents audio and video from being copied as it travels over connections.
    "cplspcon",

    # Microsoft gaming services. Auto-restarts after 1 minute.
    #"GamingServices",

    # Microsoft gaming services.
    "GamingServicesNet",

    "HP` Comm` Recover",

    # Auto-restarts after 1 day.
    "HPAppHelperCap",

    # Auto-restarts after 1 day.
    "HPDiagsCap",

    # Auto-restarts after 1 day.
    "HPNetworkCap",

    # Auto-restarts after 1 day.
    "HPOmenCap",

    # Auto-restarts after 1 day.
    "HPPrintScanDoctorService",

    # Auto-restarts after 1 day.
    "HPSysInfoCap",

    # Auto-restarts after 1 day.
    "HpTouchpointAnalyticsService",

    # Intel Bluetooth service.
    "ibtsiva",

    # Intel Graphics Command Centre service.
    "igccservice",

    # Intel HD Graphics Control Panel service.
    "igfxCUIService2.0.0.0",

    # Runs Java code in a isolated execution environment for security.
    "jhi_service",

    # Only used if an external device is connected by Thunderbolt cable.
    "TbtHostControllerService",

    # <As above>.
    "TbtP2pShortcutService",

    # Windows Search.
    "WSearch",

    # Windows Update service.
    "wuauserv"
  )

  list_service <- list(
    names = target_services,
    exists = rep(FALSE, times = length(target_services)),
    was_running = rep(FALSE, times = length(target_services))
  )

  for (i in seq_along(list_service$name)) {

    # Test if services exist via PowerShell.
    exists_test <- tryCatch(
      expr = {
        sys::exec_internal(
          "powershell",
          args = c("Get-Service -Name", list_service$names[i],
                   "-ErrorAction SilentlyContinue")
        )
      },
      error = function(e) {
        NULL
      }
    )

    if (!is.null(exists_test)) {
      list_service$exists[i] <- TRUE

      # Test if services are running.
      running_test <- tryCatch(
        expr = {
          sys::as_text(
            sys::exec_internal(
              "powershell",
              args = c(
                "Get-Service -Name", list_service$names[i],
                "| Where-Object {$_.Status -eq 'Running'} -ErrorAction SilentlyContinue"
              )
            )$stdout
          )
        },
        error = function(e) {
          NULL
        }
      )

      if (!is.null(running_test)) {
        if (length(running_test) > 0) {
          list_service$was_running[i] <- TRUE

          # Build service names string.
          if (is.null(consolidated_service_names)) {
            consolidated_service_names <- list_service$names[i]
          } else {
            consolidated_service_names <- paste0(
              consolidated_service_names, ", ", list_service$names[i]
            )
          }

        }
      }
    }
  }

  # Close relevant services.
  if (!is.null(consolidated_service_names)) {
    tryCatch(
      expr = {
        sys::exec_wait(
          "powershell",
          args = paste0("Stop-Service -Name ", consolidated_service_names),
          std_err = FALSE
        )
        # Alternative if running without elevated rights. Currently disabled.
        # sys::exec_wait(
        #   "powershell",
        #   args = paste0(
        #     "Start-Process PowerShell -ArgumentList '-Command & {Stop-Service -Name ",
        #     consolidated_service_names, "}' -Verb RunAs -WindowStyle hidden"
        #   ),
        #   std_err = FALSE
        # )
      },
      error = function(e) {
        NULL
      }
    )
  }

  # Wait 90 seconds. Stopping "igfxCUIService2.0.0.0" triggers a child process
  # which runs for about 60-90 seconds.
  if (list_service$was_running[which(list_service$names == "igfxCUIService2.0.0.0")] == TRUE) {
    Sys.sleep(90)
  }


  print("Services and processes disabled.")

  # Return value...
  list(services = list_service, processes = list_process)
}


background_services_enable <- function(list_restart) {

  # Dynamically re-enable system services and processes.
  #
  # Input:
  #   list_restart    list(services = list_service, processes = list_process)
  #
  #   where,
  #   list_service    a list describing services that were shut down.
  #   list_process    a list describing processes that were shut down.
  #
  # It's assumed list_restart was generated by a call to
  # background_services_disable().
  #
  #   N.B. If the function doesn't complete, NULL is returned.


  # Start services section #####################################################


  print("Re-enabling services...")

  # Ensure the session is being run with administrator privileges.
  # N.B. Test for text "True" / "False" here, not Boolean TRUE / FALSE.
  if (sys::as_text(sys::exec_internal("powershell", args = "[Security.Principal.WindowsIdentity]::GetCurrent().Groups -contains 'S-1-5-32-544'")$stdout) != "True") {
    cat("Function cancelled. Should not be executed without elevated rights.")
    return(NULL)
  }


  consolidated_service_names <- NULL

  for (i in seq_along(list_restart$services$names)) {
    if (list_restart$services$was_running[i] == TRUE) {

      # Build service names string.
      if (is.null(consolidated_service_names)) {
        consolidated_service_names <- list_restart$services$name[i]
      } else {
        consolidated_service_names <- paste0(
          consolidated_service_names, ", ", list_restart$services$names[i]
        )
      }
    }
  }

  # Start relevant services.
  if (!is.null(consolidated_service_names)) {
    sys::exec_wait("powershell", args = paste0("Start-Service -Name ", consolidated_service_names))
    # Alternative for sessions not being run with elevated rights. Currently disabled.
    #sys::exec_wait("powershell", args = paste0("Start-Process PowerShell -ArgumentList '-Command & {Start-Service -Name ", consolidated_service_names, "}' -Verb RunAs -WindowStyle hidden"))
  }


  # Start processes section ####################################################

  print("Re-enabling processes...")

  consolidated_commands <- NULL

  path_index <- 0
  for (i in seq_along(list_restart$processes$names)) {

    # Don't restart a process that hasn't been confirmed as present.
    if (list_restart$processes$exists[i] == FALSE) {
      next
    }

    # Don't restart a process that was previously suspended.
    if (list_restart$processes$was_suspended[i] == TRUE) {
      next
    }

    # Don't start a second process instance if one is already running.
    # (Some explicitly started processes may themselves trigger startup
    #  of other processes in the target list here.)
    # So, test if processes are running via PowerShell.
    path_index <- path_index + 1
    exists_test <- tryCatch(
      expr = {
        sys::exec_internal(
          "powershell",
          args = paste0(
            "Get-Process '",
            list_restart$processes$names[i],
            "' -ErrorAction SilentlyContinue"
          )
        )
      },
      error = function(e) {
        NULL
      }
    )
    if (!is.null(exists_test)) {
      next
    }

    # Build process names string.
    if (is.null(consolidated_commands)) {
      consolidated_commands <- paste0(
        "Start-Process -WindowStyle hidden -FilePath \"",
        gsub(" ", "` ", list_restart$processes$paths[path_index]),
        "\""
      )
    } else {
      consolidated_commands <- paste0(
        consolidated_commands,
        "; Start-Process -WindowStyle hidden -FilePath \"",
        gsub(" ", "` ", list_restart$processes$paths[path_index]),
        "\""
      )
    }


  }

  # Start relevant processes.
  if (!is.null(consolidated_commands)) {
    sys::exec_wait("powershell", args = consolidated_commands)
    # Alternative for sessions not being run with elevated rights. Currently disabled.
    #sys::exec_wait("powershell", args = paste0("Start-Process PowerShell -ArgumentList '-Command & {", consolidated_commands, "}' -Verb RunAs -WindowStyle hidden"))
  }


  # Restart the Task Scheduler.
  # Modify the registry key: HKEY_LOCAL_MACHINE\SYSTEM\CurrentControlSet\Services\Schedule.
  # N.B. Change the key's Start value.
  #         (Here 2 is an alias for "Automatic (Delayed Start)")
  #         (Here 4 is an alias for "Disabled)
  print("Re-enabling Task Scheduler...")

  # Does the registry key need updating?
  key_value <- sys::as_text(sys::exec_internal("powershell",
    args = "$key_val = Get-ItemProperty -Path 'HKLM:\\SYSTEM\\CurrentControlSet\\Services\\Schedule'; $key_val.Start")$stdout)
  if (as.integer(key_value) == 4) {
    sys::exec_wait("powershell", args = "Set-ItemProperty -Path 'HKLM:\\SYSTEM\\CurrentControlSet\\Services\\Schedule' -Name 'Start' -Value 2")
    cat("Task Scheduler registry details updated. Please restart system.\n")
    return(NULL)
  }

  # Is the service still disabled?
  running_test <- tryCatch(
    expr = {
      sys::as_text(sys::exec_internal("powershell", args = c("Get-Service -Name 'Schedule' | Where-Object {$_.Status -eq 'Running'}"))$stdout)
    },
    error = function(e) {NULL}
  )
  if (is.null(running_test) | (length(running_test) == 0)) {
    cat("Task Scheduler set to start. Please restart system.\n")
    return(NULL)
  }


  # Return value...
  "Services and processes enabled."
}


calc_mcmc_runtimes <- function(input_model, input_data, active_cores,
                               input_seq, hyperthread) {

  # Get runtimes for repeated execution of an input Stan model.
  #
  # Parameters:
  #   input_model     A Stan model.
  #   input_data      Data to pass to the Stan model.
  #   active_cores    An integer specifying the number of cores to use for Stan.
  #   input_seq       A sequence of iterations to test, e.g. 1 .. 200.
  #   hyperthread     A boolean specifying whether to use hyperthreading.
  #
  # Output:
  #   A data frame of runtimes for each iteration.

  # Load legacy results, if no model supplied.
  if (is.null(input_model)) {
    load(file = "adass_model.rda")
  }

  vec_runtimes <- rep(NA, length(input_seq))

  # Here a 'sampling' call is used with a compiled stan model (produced earlier
  # with 'stan_model') instead of using a 'stan' call. Maintaining an explicit
  # reference to a compiled stan model stops it getting garbage collected, in
  # turn preventing "recompiling to avoid crashing R session" operations. If
  # a single 'stan' call was used here instead, the compiled model would be
  # eligible for garbage collection after each evaluation of microbenchmark.
  #
  # N.B. The 'rstan_options(auto_write = TRUE)' call during setup only prevents
  # recompilation of unchanged AND un-garbage collected stan models.

  # Set default benchmarking time unit to seconds.
  # Remember, microbenchmark does (needed) warmup iterations.


  # Collect runtimes.
  iteration <- 0
  for (i in input_seq) {

    iteration <- iteration + 1

    # Give a progress indication.
    cat("get_mcmc_median_runtimes progress: evaluating cores = ", active_cores,
      ", chains = ", active_cores, "...\n", sep = "")

    vec_runtimes[iteration] <- summary(
      microbenchmark::microbenchmark(
        sampling(
          input_model,
          data = input_data,
          chains = active_cores,
          warmup = 340,   # Stan default is iter/2.
          iter = 680,     # 700 for physical cores, 680 logical.
          seed = 1
        ),
        times = 1,
        unit = "s"
      )
    )$median

    # Print results in units of seconds.
    cat(
      "# cores = ", active_cores,
      ", # chains = ", active_cores,
      ", median runtime = ", vec_runtimes[iteration], " seconds.",
      " (iteration = ", iteration, ")\n",
      sep = ""
    )

  }

  # Return results, packaged into a data frame...
  data.frame(
    iterations = input_seq,
    runtimes = vec_runtimes,
    dummy = factor(0)
  )
}


plot_finalise <- function(target_file) {

  # Close the active plot and embed fonts from the current font map into it.
  #
  # Parameters:
  #   target_file   Character; a target file name.
  #
  # Examples usage:
  #   plot_finalise("test_file.pdf")

  # Only proceed if the specified file exists.
  if (file.exists(target_file)) {

    # Shut down the graphics device used to render the plot.
    # Don't try and shut down the null device (device 1), if it's active.
    while (!is.null(grDevices::dev.list())) {
      device_num <- as.integer(dev.cur())
      if (device_num != 1) {
        dev.off(which = device_num)
      }

      # Release memory.
      rm(device_num)
    }

    embed_fonts(target_file)
    call(plot_crop(target_file))
  }
}


plot_fix_panels <- function(input_plots, scale_x = 1, scale_y = 1) {

  # Apply physical dimensions to a plot panel area.
  #
  # Parameters:
  #   input_plots   One or more ggplot2::ggplot plots.
  #   scale_x       A numeric value: a plot width scaling factor.
  #   scale_y       A numeric value: a plot height scaling factor.
  #
  # Example usage:
  #    plot_fix_panels(ggplot2::ggplot() + ggplot2::theme_void())

  panel_side_length <- 42   # 20% of A4 width = 0.2 * 210 mm.

  lapply(
    list(input_plots),
    set_panel_size,
    width = unit(panel_side_length * scale_x, "mm"),
    height = unit(panel_side_length * scale_y, "mm")
  )
}


plot_mcmc_runtimes <- function(input_dataframe, plot_filename) {

  # Plot runtimes for repeated execution of an input Stan model.
  #
  # Parameters:
  #   input_dataframe   Dataframe holding benchmarking results.
  #   plot_filename     A file name to save the plot as.

  # Plot runtimes vs iterations.
  pdf(plot_filename)
  #
  xbreaks <- seq(0, nrow(input_dataframe), by = 50)
  xlimits <- c(0, nrow(input_dataframe))
  ybreaks <- seq(0, 88, 22)   # Determined empirically.
  ylabel <- 5.50
  ylimits <- c(0, 88)
  #
  # Generate plot.
  ggplot_runtimes <- ggplot(
    input_dataframe, aes(x = iterations, y = runtimes)
  ) +
    geom_point(aes(shape = dummy), size = 1) +
    geom_line(aes(group = dummy), size = 0.25) +
    labs(x = "Iteration",
         y = TeX("Runtime $\\,$(sec)")) +
    scale_shape_manual(values = c(shape_round)) +
    scale_x_continuous(breaks = xbreaks, limits = xlimits) +
    scale_y_continuous(breaks = ybreaks, limits = ylimits) +
    theme_throttlestop() +
    geom_label(
      label = TeX(
        paste0(
          "variance$ \\approx ",
          format(round(var(input_dataframe$runtimes), digits = 1), nsmall = 1),
          "$\\,sec"
        ),
        output = "character"
      ),
      family = "CM Roman",
      parse = TRUE,
      size = 2.8,
      x = 136,
      y = ylabel
    )
  grid.arrange(grobs = plot_fix_panels(ggplot_runtimes))
  plot_finalise(plot_filename)

  # Return value...
  NULL
}


theme_throttlestop <- function() {

  # A custom ggplot theme.
  # Returns a ggplot2::theme.
  #
  # Example usage:
  #   plot_temp <- ggplot2::ggplot() + theme_throttlestop()

  # Ensure required fonts are available.
  stopifnot(validate_ghostscript_paths())
  loadfonts(quiet = TRUE)

  ggplot2::theme(
    axis.line = element_line(size = 0.4),
    axis.text = element_text(colour = "black"),
    axis.ticks = element_line(colour = "black"),
    legend.position = "none",
    panel.background = element_rect(fill = "white"),
    panel.border = element_rect(colour = "black", fill = "transparent"),
    panel.grid.major = element_line(
      size = 0.25,
      linetype = "dashed",
      colour = "grey"
    ),
    panel.grid.minor = element_line(
      size = 0.25,
      linetype = "dashed",
      colour = "grey"
    ),
    text = element_text(family = "CM Roman", size = 10)
  )
}


validate_ghostscript_paths <- function() {

  # Return a logical value (TRUE or FALSE) indicating whether the expected
  # Ghostscript-related environment variables have been set to valid values.


  # Validate single-valued environment variables.
  ghostscript_env_vars <- c("GS_CMD", "R_GSCMD")
  for (env_var in ghostscript_env_vars) {
    path_test <- Sys.getenv(env_var)
    if (nchar(path_test) == 0) {
      cat("Environment variable ", env_var, " is missing.\n", sep = "")
      return(FALSE)

    } else if (!file.exists(path_test)) {
      cat("Environment variable ", env_var, " holds an invalid path.\n",
          sep = "")
      return(FALSE)
    }
  }


  # Validate multi-valued environment variables.
  env_var <- "GS_FONTPATH"
  multi_path_test <- Sys.getenv(env_var)
  if (nchar(multi_path_test) == 0) {
    cat("Environment variable ", env_var, " is missing.\n", sep = "")
    return(FALSE)
  }

  for (path_test in strsplit(multi_path_test, ";")[[1]]) {
    if (nchar(path_test) == 0) {
      cat("Environment variable ", env_var, " is malformed.\n", sep = "")
      return(FALSE)

    } else if (!file.exists(path_test)) {
      cat("Environment variable ", env_var, " holds an invalid path.\n",
          sep = "")
      return(FALSE)
    }
  }

  # Return value...
  TRUE
}


# Font config.
stopifnot(validate_ghostscript_paths())
loadfonts(quiet = TRUE)

```


```{r mcmc_sampling}

# From 'Bayesian Models for Astrophysical Data',
# by Hilbe, de Souza & Ishida, 2016, Cambridge Univ Press.
#
# Code 10.26 Bayesian normal model for cosmological parameter
#            inference from type Ia supernova data in R using Stan.
#
# Statistical Model: Gaussian regression in R using Stan.
#                    Example using ODE.
#
# Astronomy case: Cosmological parameters inference from
#                 type Ia supernovae data.
#
# Data: JLA sample, Betoule et al., 2014
# http://supernovae.in2p3.fr/sdss_snls_jla/ReadMe.html
#
# 1 response (obsy - observed magnitude)
# 5 explanatory variable (redshift - redshift,
#                         ObsMag   - apparent magnitude,
#                         x1       - stretch,
#                         color    - colour,
#                         hmass    - host mass)


# Set initial conditions.
z0 <- 0                         # Initial redshift.
E0 <- 0                         # Integral(1/E) at z0.

# Physical constants.
c <- 3e5                        # Speed of light (km).
H0 <- 70                        # Hubble constant.

# Import data.
# Remote source:
data <- read.table(
  "https://raw.githubusercontent.com/astrobayes/BMAD/master/data/Section_10p11/jla_lcparams.txt",
  header = TRUE
)
# Local source, if available:
#data <- read.table("jla_lcparams.txt", header = TRUE)

# Remove repeated redshifts.
data <- data[!duplicated(data$zcmb), ]

# Prepare data for Stan.
index    <- order(data$zcmb)                     # Sort according to redshift.
stan_data <- list(nobs = nrow(data),             # Number of SNe. Expect 732.
                  E0 = array(E0, dim = 1),
                  z0 = z0,
                  c = c,
                  H0 = H0,
                  obs_mag = data$mb[index],      # Apparent magnitude.
                  redshift = data$zcmb[index],   # Redshift.
                  x1 = data$x1[index],           # Stretch.
                  color = data$color[index],     # Colour.
                  hmass = data$m3rdvar[index])   # Host mass.

# Release unneeded memory.
rm(c)
rm(data)
rm(E0)
rm(H0)
rm(index)
rm(z0)

# Fit
stan_model <- "
functions {
    // ODE for the inverse Hubble parameter.
    // System State E is 1 dimensional.
    // The system has 2 parameters theta = (om, w)
    //
    // where
    //
    //   om:   dark matter energy density
    //   w:    dark energy equation of state parameter
    //
    // The system redshift derivative is
    //
    // d.E[1] / d.z = 1.0/sqrt(om * pow(1+z,3) + (1-om) * (1+z)^(3 * (1+w)))
    //
    // @param z redshift at which derivatives are evaluated.
    // @param E system state at which derivatives are evaluated.
    // @param params parameters for system.
    // @param x_r real constants for system (empty).
    // @param x_i integer constants for system (empty).

    real[] Ez(real z,
              real[] H,
              real[] params,
              real[] x_r,
              int[] x_i) {
       real dEdz[1];
       dEdz[1] = 1.0 / sqrt(params[1] * (1 + z)^3
                 + (1 - params[1]) * (1 + z)^(3 * (1 + params[2])));
       return dEdz;
    }
}
data {
   int<lower=1> nobs;            // number of data points
   real E0[1];                   // integral(1/H) at z=0  
   real z0;                      // initial redshift, 0
   real c;                       // speed of light
   real H0;                      // Hubble parameter
   vector[nobs] obs_mag;         // observed magnitude at B max
   real x1[nobs];                // stretch
   real color[nobs];             // colour
   real redshift[nobs];          // redshift
   real hmass[nobs];             // host mass
}
transformed data {
   real x_r[0];                  // required by ODE (empty)
   int x_i[0]; 
}
parameters{
   real<lower=0, upper=1> om;    // dark matter energy density
   real alpha;                   // stretch coefficient   
   real beta;                    // color coefficient
   real Mint;                    // intrinsic magnitude
   real deltaM;
   real<lower=0> sigint;         // magnitude dispersion
   real<lower=-2, upper=0> w;    // dark matter equation of state parameter
}
transformed parameters{
   real DC[nobs,1];              // co-moving distance 
   real pars[2];                 // ODE input = (om, w)
   vector[nobs] mag;             // apparent magnitude
   real dl[nobs];                // luminosity distance
   real DH;                      // Hubble distance = c/H0

   DH = c / H0;
   pars[1] = om;
   pars[2] = w;

   // Integral of 1/E(z)
   DC = integrate_ode_rk45(Ez, E0, z0, redshift, pars,  x_r, x_i);
   for (i in 1:nobs) {
      dl[i] = DH * (1 + redshift[i]) * DC[i, 1];
      if (hmass[i] < 10) mag[i] = 25 + 5 * log10(dl[i]) + Mint - alpha * x1[i] + beta * color[i];
      else mag[i] = 25 + 5 * log10(dl[i]) + Mint + deltaM - alpha * x1[i] + beta * color[i];
   }
}
model {
   // Priors and likelihood.
   sigint ~ gamma(0.001, 0.001);
   Mint ~ normal(-20, 5.);
   beta ~ normal(0, 10);
   alpha ~ normal(0, 1);
   deltaM ~ normal(0, 1);
   obs_mag ~ normal(mag, sigint);
}
"

```


```{r test_convergence}

# Generate and save the Stan model.
stan_model_eval <- stan_model(model_code = stan_model, save_dso = FALSE)
save(stan_model_eval, file = paste0("adass_model.rda"))

# Test MCMC parameters are consistent with convergence.
fit <- sampling(
  stan_model_eval,
  data = stan_data,
  chains = active_cores,
  warmup = 340,   # 350 for 6 physical cores, 340 for 12 logical.
  iter = 680,     # 700 for 6 physical cores, 680 for 12 logical.
  seed = 1
)

# Output results.
print(fit, pars = c("om", "Mint", "w", "alpha", "beta", "deltaM", "sigint"),
      intervals = c(0.025, 0.975), digits = 3)
print(summarise_draws(fit))   # Posterior package results.
#
# a) Results are normally ok for Rhat <= 1.05 (from the vanilla rstan results).
#    Ideally however, Rhat < 1.01 is sought (from the posterior package).
#    - The posterior package uses an improved version of the traditional Rhat.
#    - To get high precision values, use: summarise_draws(fit_synth)$rhat
#
# b) Results are normally ok if n_eff > 100.
#    Ideally however, a ratio of n_eff / N >= 0.1 is sought.
#    (where N = total samples.)
#    Ref: https://mc-stan.org/bayesplot/articles/visual-mcmc-diagnostics.html

# Free memory.
rm(fit)
gc()
```


```{r dampen_os_noise}

# Disable selected background processes and services.
tmp <- background_services_disable()
stopifnot(!is.null(tmp))

# It's assumed this code will be quit and Windows restarted after toggling the
# active status of the Task Manager.
# It's also assumed the RStudio session will be saved (and later reloaded) prior
# to continuing from this point. See too the header notes in this Rmd file.

```


```{r test_mcmc_throttled}

# Evaluate typical performance of parallelised MCMC.

# Ensure ThrottleStop is running.
stopifnot(
  grepl(
    "ThrottleStop.exe",
    shell(
      'tasklist /fi "imagename eq ThrottleStop.exe" /nh /fo csv',
      intern = TRUE
    )
  )
)

# Collect and plot MCMC runtimes, throttled.
# It's assumed the "Disable Turbo" checkbox in ThrottleStop is checked at this
# point.
iterations_lbound <- 1
iterations_ubound <- 200
iterations_seq <- seq(iterations_lbound, iterations_ubound, by = 1)

df_bmark_throttled <- calc_mcmc_runtimes(
  input_model = stan_model_eval,
  input_data = stan_data,
  active_cores = active_cores,
  input_seq = iterations_seq,
  hyperthread = hyperthreading_status
)
plot_mcmc_runtimes(
  input_dataframe = df_bmark_throttled,
  plot_filename = "plot_runtimes_vs_iterations_throttled.pdf"
)

```


```{r test_mcmc_unthrottled}

# It's assumed turbo is re-enabled and ThrottleStop is closed at this point.
# Ensure ThrottleStop is closed.
stopifnot(
  !grepl(
    "ThrottleStop.exe",
    shell(
      'tasklist /fi "imagename eq ThrottleStop.exe" /nh /fo csv',
      intern = TRUE
    )
  )
)

# Collect and plot MCMC runtimes, unthrottled.
df_bmark_unthrottled <- calc_mcmc_runtimes(
  input_model = stan_model_eval,
  input_data = stan_data,
  active_cores = active_cores,
  input_seq = iterations_seq,
  hyperthread = hyperthreading_status
)
plot_mcmc_runtimes(
  input_dataframe = df_bmark_unthrottled,
  plot_filename = "plot_runtimes_vs_iterations_unthrottled.pdf"
)

# Re-enable selected background processes and services.
stopifnot(!is.null(background_services_enable(tmp)))
rm(tmp)
```
